---
title: "NM2207- Final Submission"
---

# San Francisco's Changing Tree Landscape: A Decade of Growth and Diversity

Have you heard of **San Francisco Trees?!** Let **ME**.... **TELL YOU** more about it!

## How has the tree population in San Francisco changed over the past decade?

> According to studies on urban forestry and sustainability, the state of a city's tree population can significantly impact the environment, air quality, and overall well-being of its residents. For instance, research by the United Nations has shown that urban green spaces, including trees, are vital for mitigating climate change. This question is important because understanding how San Francisco's tree population has evolved is crucial for assessing the city's progress in environmental sustainability and urban planning, which has wide-reaching effects on residents' quality of life.

## Dataset

```{r, eval=TRUE,echo=FALSE}

library(tidyverse)
library(ggmap)
library(osmdata)
library(extrafont)
library(sf)
library(lubridate)
library(viridis)
library(patchwork)
library(ggtext)

loadfonts(device = "win")

sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

roads <- getbb("San Francisco") %>% 
  opq() %>% 
  add_osm_feature("highway") %>% 
  osmdata_sf()

coast <- getbb("San Francisco") %>% 
  opq() %>% 
  add_osm_feature(key = "natural", value = "coastline") %>%
  osmdata_sf()

sf_trees %>%
  summarise(nb_species = length(unique(species)))

# Vector of years to fill the gap in the dataset

years <- tibble(year = seq(1954,2019,1), nb = 0)

# Number of different species per year

p1 <- sf_trees %>%
  mutate(year = year(date)) %>%
  filter(year != 2020) %>%
  group_by(year) %>%
  summarise(nb_species = length(unique(species))) %>%
  right_join(years) %>%
  mutate(nb_species = ifelse(is.na(nb_species), 0, nb_species)) %>% 
  ggplot(aes(year, nb_species)) +
  geom_col(fill = "#6DD19C", color = "#effdee", size = 1) +
  scale_x_continuous(limits=c(1955,2020))+
  labs(title= "Number of different specie planted each year",
       y = "# different specie") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white", color = NA),
        legend.background = element_rect(fill = "white"),
        plot.title = element_text(color = "#555555", hjust = 0.5),
        panel.grid.major =  element_line(color = "#555555", linetype = "dashed"),
        panel.grid.minor =  element_line(color = "#555555", linetype = "dotted"),
        axis.text = element_text(color = "#555555"),
        axis.title = element_text(color = "#555555"),
        axis.title.x = element_blank())


# Number of trees planted each year

p2 <- sf_trees %>%
  mutate(year = year(date)) %>%
  filter(year != 2020 ) %>%
  filter(!is.na(year)) %>%
  group_by(year) %>%
  summarise(nb_trees = n()) %>% 
  right_join(years) %>%
  mutate(nb_trees = ifelse(is.na(nb_trees), 0, nb_trees)) %>%
  ggplot(aes(year, nb_trees)) +
  geom_line(color = "#6DD19C", size = 1) +
  scale_x_continuous(limits=c(1955,2020))+
  labs(title= "Number of trees planted each year",
       y = "number of trees") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white", color = NA),
        legend.background = element_rect(fill = "white"),
        plot.title = element_text(color = "#555555", hjust = 0.5),
        panel.grid.major =  element_line(color = "#555555", linetype = "dashed"),
        panel.grid.minor =  element_line(color = "#555555", linetype = "dotted"),
        axis.text = element_text(color = "#555555"),
        axis.title = element_text(color = "#555555"),
        axis.title.x = element_blank())


patchwork <- ( p1 / p2 )

patchwork + plot_annotation(title = "San Francisco Trees",
                            subtitle = "Evolution of the number of trees & species.",
                            theme = theme(plot.background = element_rect(fill = "white"),
                                        plot.title = element_text(color = "#555555", size = 18, hjust = 0.5),
                                        plot.subtitle = element_text(color = "#555555", size = 14, hjust = 0.5),
                                        plot.caption = element_text(color = "#555555", hjust = 1),
                                        panel.background = element_rect(fill = "white")))+
            plot_layout(nrow = 2, heights = c(0.05, 5))


```


```{r, eval=TRUE,echo=FALSE}
library(tidyverse)
library(sf)

theme_set(theme_light())

sf_trees_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

sf_trees <- sf_trees_raw %>%
  # 1. Remove outliers
  filter(longitude > -130,
         between(latitude, 37.6, 40)) %>%
  # 2. Make missing values explicit
  mutate(species = ifelse(species %in% c("Tree(s) ::", "::"), NA_character_, species)) %>%
  # 3. Separate species into scientific and common names
  separate(species,
           into = c("science_name", "common_name"),
           sep = "::") %>%
  # Get rid of whitespace that occurs after separation
  mutate_at(vars(science_name, common_name), str_squish) %>%
  # 4. Use scientific name for missing common names
  mutate(common_name = ifelse(common_name == "", science_name, common_name))

# 1. Create a temp directory and download ZIP file from US Census
temp_download <- tempfile()
download.file("https://www2.census.gov/geo/tiger/TIGER2017//ROADS/tl_2017_06075_roads.zip", temp_download)

# 2. Create a temp directory to unzip the file into
temp_unzip <- tempfile()
unzip(temp_download, exdir = temp_unzip)

# 3. Read the unzipped shapefile from the temporary filepath
sf_roads <- read_sf(paste0(file.path(temp_unzip), "/tl_2017_06075_roads.shp"))

sf_trees %>%
  mutate(
    common_name = ifelse(
      common_name == "Indian Laurel Fig Tree 'Green Gem'",
      "Indian Laurel Fig",
      common_name
    ),
    common_name = fct_lump(common_name, 9)
  ) %>%
  filter(!is.na(common_name), common_name != "Other") %>%
  add_count(common_name, name = "tree_count") %>%
  mutate(common_name = fct_reorder(common_name,-tree_count)) %>%
  ggplot() +
  geom_sf(data = sf_roads, col = "grey90", alpha = 0.5) +
  geom_point(
    aes(longitude, latitude),
    shape = ".",
    alpha = 0.5,
    col = "#6DD19C"
  ) +
  facet_wrap( ~ common_name) +
  coord_sf() +
  labs(title = "Where are San Francisco's trees?",
       subtitle = "Showing the nine most common tree species") +
  theme_void() +
  theme(
    plot.subtitle = element_text(margin = margin(b = 4, unit = "pt")),
    strip.text = element_text(size = 8, margin = margin(
      t = 2, b = 2, unit = "pt"
    ))
  )

```

