---
title: "NM2207- Final Submission"
---

# San Francisco's Changing Tree Landscape: A Decade of Growth and Diversity

Have you heard of **San Francisco Trees?!** Let **ME**.... **TELL YOU** more about it!

## How has the tree population in San Francisco changed over the past decade?

> According to studies on urban forestry and sustainability, the state of a city's tree population can significantly impact the environment, air quality, and overall well-being of its residents. For instance, research by the United Nations has shown that urban green spaces, including trees, are vital for mitigating climate change. This question is important because understanding how San Francisco's tree population has evolved is crucial for assessing the city's progress in environmental sustainability and urban planning, which has wide-reaching effects on residents' quality of life.

# Why is it important to address this question?

> It is important to address this question because the condition and evolution of a city's tree population have direct implications for environmental sustainability, air quality, and residents' quality of life. Understanding these changes allows for informed urban planning, conservation efforts, and assessing progress toward climate and environmental goals.

## Findings about San Francisco Trees

# Graphs on the trees populations and planting efforts

```{r, eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}

library(tidyverse)
library(ggmap)
library(osmdata)
library(extrafont)
library(sf)
library(lubridate)
library(viridis)
library(patchwork)
library(ggtext)

loadfonts(device = "win")

sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

roads <- getbb("San Francisco") %>% 
  opq() %>% 
  add_osm_feature("highway") %>% 
  osmdata_sf()

coast <- getbb("San Francisco") %>% 
  opq() %>% 
  add_osm_feature(key = "natural", value = "coastline") %>%
  osmdata_sf()

sf_trees %>%
  summarise(nb_species = length(unique(species)))

# Vector of years to fill the gap in the dataset

years <- tibble(year = seq(1954,2019,1), nb = 0)

# Number of different species per year

p1 <- sf_trees %>%
  mutate(year = year(date)) %>%
  filter(year != 2020) %>%
  group_by(year) %>%
  summarise(nb_species = length(unique(species))) %>%
  right_join(years) %>%
  mutate(nb_species = ifelse(is.na(nb_species), 0, nb_species)) %>% 
  ggplot(aes(year, nb_species)) +
  geom_col(fill = "#6DD19C", color = "#effdee", size = 1) +
  scale_x_continuous(limits=c(1955,2020))+
  labs(title= "Number of different specie planted each year",
       y = "# different specie") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white", color = NA),
        legend.background = element_rect(fill = "white"),
        plot.title = element_text(color = "#555555", hjust = 0.5),
        panel.grid.major =  element_line(color = "#555555", linetype = "dashed"),
        panel.grid.minor =  element_line(color = "#555555", linetype = "dotted"),
        axis.text = element_text(color = "#555555"),
        axis.title = element_text(color = "#555555"),
        axis.title.x = element_blank())


# Number of trees planted each year

p2 <- sf_trees %>%
  mutate(year = year(date)) %>%
  filter(year != 2020 ) %>%
  filter(!is.na(year)) %>%
  group_by(year) %>%
  summarise(nb_trees = n()) %>% 
  right_join(years) %>%
  mutate(nb_trees = ifelse(is.na(nb_trees), 0, nb_trees)) %>%
  ggplot(aes(year, nb_trees)) +
  geom_line(color = "#6DD19C", size = 1) +
  scale_x_continuous(limits=c(1955,2020))+
  labs(title= "Number of trees planted each year",
       y = "number of trees") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white", color = NA),
        legend.background = element_rect(fill = "white"),
        plot.title = element_text(color = "#555555", hjust = 0.5),
        panel.grid.major =  element_line(color = "#555555", linetype = "dashed"),
        panel.grid.minor =  element_line(color = "#555555", linetype = "dotted"),
        axis.text = element_text(color = "#555555"),
        axis.title = element_text(color = "#555555"),
        axis.title.x = element_blank())


patchwork <- ( p1 | p2 )

patchwork + plot_annotation(title = "Evolution of the number of trees & species",
                            theme = theme(plot.background = element_rect(fill = "white"),
                                        plot.title = element_text(color = "#555555", size = 18, hjust = 0.5),
                                        plot.subtitle = element_text(color = "#555555", size = 14, hjust = 0.5),
                                        plot.caption = element_text(color = "#555555", hjust = 1),
                                        panel.background = element_rect(fill = "white")))+
            plot_layout(nrow = 2, heights = c(10, 10))


```

> In graph above, the number of trees planted in San Francisco kept increasing from the [70's]{style="color:#feb24c"} to the end of the [90's]{style="color:#feb24c"}. This period showed an increase of the number of different species planted each year, with less than [50]{style="color:#00d400"} to about [150]{style="color:#00d400"}, with the maximum of [173 different species]{style="color:#00d400"} in [2008]{style="color:#feb24c"}, and [4388 trees planted]{style="color:#00d400"}. The top graph shows the evolution of the number of species planted each year, while the bottom one shows the evolution of the number of trees planted each year.

# Identifying the most common species provides valuable information on the dominant types of trees in San Francisco

```{r, eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(sf)

theme_set(theme_light())

sf_trees_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

sf_trees <- sf_trees_raw %>%
  # 1. Remove outliers
  filter(longitude > -130,
         between(latitude, 37.6, 40)) %>%
  # 2. Make missing values explicit
  mutate(species = ifelse(species %in% c("Tree(s) ::", "::"), NA_character_, species)) %>%
  # 3. Separate species into scientific and common names
  separate(species,
           into = c("science_name", "common_name"),
           sep = "::") %>%
  # Get rid of whitespace that occurs after separation
  mutate_at(vars(science_name, common_name), str_squish) %>%
  # 4. Use scientific name for missing common names
  mutate(common_name = ifelse(common_name == "", science_name, common_name))

# 1. Create a temp directory and download ZIP file from US Census
temp_download <- tempfile()
download.file("https://www2.census.gov/geo/tiger/TIGER2017//ROADS/tl_2017_06075_roads.zip", temp_download)

# 2. Create a temp directory to unzip the file into
temp_unzip <- tempfile()
unzip(temp_download, exdir = temp_unzip)

# 3. Read the unzipped shapefile from the temporary filepath
sf_roads <- read_sf(paste0(file.path(temp_unzip), "/tl_2017_06075_roads.shp"))

sf_trees %>%
  mutate(
    common_name = ifelse(
      common_name == "Indian Laurel Fig Tree 'Green Gem'",
      "Indian Laurel Fig",
      common_name
    ),
    common_name = fct_lump(common_name, 9)
  ) %>%
  filter(!is.na(common_name), common_name != "Other") %>%
  add_count(common_name, name = "tree_count") %>%
  mutate(common_name = fct_reorder(common_name,-tree_count)) %>%
  ggplot() +
  geom_sf(data = sf_roads, col = "grey90", alpha = 0.5) +
  geom_point(
    aes(longitude, latitude),
    shape = ".",
    alpha = 0.5,
    col = "#6DD19C"
  ) +
  facet_wrap( ~ common_name) +
  coord_sf() +
  labs(title = "The nine most common tree species") +
  theme_void() +
  theme(
    plot.subtitle = element_text(margin = margin(b = 4, unit = "pt")),
    strip.text = element_text(size = 8, margin = margin(
      t = 8, b = 8, unit = "pt"
    ))
  )

dev.new(width=20, height=20, unit="in")
plot

```

> This graph, analyzing the most common tree species in San Francisco can provide insights into the health, resilience, and management of the urban forest, informing decisions on future planting initiatives and conservation efforts.

## San Franciso's Tree Mapping

<iframe height="600" width="100%" frameborder="no" align-items="center" src="https://18en4g-chua-chua.shinyapps.io/interactivegraph/">

<iframe>
